---
title: "Joseph-J-hwk2-2"
author: "Justin Joseph"
date: "2023-02-12"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stringr, readxl, data.table, 
               gdata, MatchIt, cobalt, Matching)

final.hcris.data=read_rds('data/output/HCRIS_Data.rds')
final.hcris=read_rds('data/output/HCRIS_Data_11.rds')
```

#Answers

### 1
```{r, echo = FALSE, message = FALSE, warning = FALSE}
#1. How many hospitals filed more than one report in the same year

# Group the data by hospital and year, and count the number of reports filed by each hospital in each year
grouped_data <- final.hcris %>% 
  group_by(provider_number, fyear) %>% 
  summarize(count= n()) 

# Filter the grouped data to only include hospitals that filed more than one report in a given year

filtered_data <- grouped_data %>%
  filter(count>1)



table1<-table(filtered_data$fyear, filtered_data$count)
table1_df<-as.data.frame.matrix(table1)
table1_df<-mutate(table1_df, countSum= table1_df$`2` + table1_df$`3`)
table1_df<-rownames_to_column(table1_df, 'year')


ggplot(table1_df, aes(x=year, y=countSum, group=1)) +
  geom_line() +
  ggtitle("Number of Hospitals with more than one report ")+
  xlab("Year") +
  ylab("Number of hospitals with more than one report from 1996 to 2017") 

```

### 2
```{r,  echo = FALSE, message = FALSE, warning = FALSE}
unique_hospitalIDS<-length(unique(final.hcris.data$provider_number))
cat("Number of unique hospital IDs", unique_hospitalIDS)

```

### 3
```{r, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(final.hcris.data, aes(x = year, y = tot_charges, group=year)) +
  geom_jitter(alpha= 0.05) +
  geom_violin(alpha=.9)

```


### 4
```{r, echo = FALSE, message = FALSE, warning = FALSE}
final.hcris.data1 <-final.hcris.data %>%
  mutate(discount_factor = 1-tot_discounts/tot_charges)%>%
  mutate(price_num = (ip_charges + icu_charges + ancillary_charges)*discount_factor - tot_mcare_payment) %>%
  mutate(price_denom = tot_discharges - mcare_discharges) %>%
  mutate(price = price_num/price_denom)

Justin <- final.hcris.data1 %>% ungroup() %>%
  filter(price_denom>100, !is.na(price_denom), 
         price_num>0, !is.na(price_num),
         price<100000, 
         beds>30) %>%
  mutate( hvbp_payment = ifelse(is.na(hvbp_payment),0,hvbp_payment),
          hrrp_payment = ifelse(is.na(hrrp_payment),0,abs(hrrp_payment)),
          penalty = (hvbp_payment-hrrp_payment<0))  
Justin1<-Justin %>%
  filter(price<25000)
ggplot(Justin1, aes(x = year, y = price)) +
  geom_jitter(alpha= 0.05) +
  geom_violin(alpha=.9)


```

### 5

```{r, echo = FALSE, message = FALSE, warning = FALSE}
avg_price_penalty <- Justin %>%
  filter(year==2012) %>%
  filter(penalty==TRUE) %>%
  summarise(penalty_price=mean(price))
avg_price_penalty  
avg_price_nopenalty <- Justin %>%
  filter(year==2012) %>%
  filter(penalty==FALSE) %>%
  summarise(no_penalty_price=mean(price))
avg_price_nopenalty  
```


### 6

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Calculate the quartiles of the data set
data_2012<-Justin%>% filter(year==2012)
quartiles <- quantile(data_2012$beds, probs = c(0.25, 0.5, 0.75,1), na.rm=TRUE)

q1 <- quartiles[1]
q2 <- quartiles[2]
q3 <- quartiles[3]
q4<- quartiles[4]


# Indicate which observations belong to which group
data_with_groups <- data_2012 %>%
  mutate(Q1 = ifelse(beds <= q1, 1, 0)) %>%
  mutate(Q2 = ifelse(beds >q1 & beds <=q2, 1, 0)) %>%
  mutate(Q3 = ifelse(beds >q2 & beds <=q3, 1, 0)) %>%
  mutate(Q4 = ifelse(beds >q3 & beds <=q4, 1, 0)) 


data_with_CT<- data_with_groups %>%
  filter(!is.na(price)) %>%
  mutate(group = ifelse(penalty==FALSE, "Control", "Treatment")) 


results<- data_with_CT %>%
  group_by(group, Q1, Q2, Q3, Q4) %>%
  summarize(mean_price=mean(price, na.rm = TRUE))

results
```

### 7

```{r, echo = FALSE, message = FALSE, warning = FALSE,}
nn.est1 <- Matching::Match(Y=data_with_CT$price,
                           Tr=data_with_CT$penalty,
                           X=(data_with_CT %>% dplyr::select(Q1, Q2, Q3, Q4)),
                           M=1,
                           Weight=1,
                           estimand="ATE")

summary(nn.est1)




nn.est2 <- Matching::Match(Y=data_with_CT$price,
                           Tr=data_with_CT$penalty,
                           X=(data_with_CT %>% dplyr::select(Q1, Q2, Q3, Q4)),
                           M=1,
                           Weight=2,
                           estimand="ATE")
summary(nn.est2)


##propensity score
logit.reg <- glm(penalty ~ Q1+ Q2+ Q3 + Q4,
                 data = data_with_CT, family = binomial)
data_with_CT <- data_with_CT %>%
  mutate(ps = predict(logit.reg, type = 'response')) %>%
  filter(ps>0 & ps<1)


# Create IPW weights
data_with_CT <- data_with_CT %>%
  mutate(ipw = case_when(
    penalty == 1 ~ 1/ps,
    penalty == 0 ~ 1/(1-ps),
    TRUE~NA_real_
  ))


mean.t1 <- data_with_CT %>% 
  filter(penalty==1) %>% 
  dplyr::select(price, ipw) %>%
  summarize(mean_y=weighted.mean(price, w=ipw))
mean.t0 <- data_with_CT %>% 
  filter(penalty==0) %>% 
  dplyr::select(price, ipw) %>%
  summarize(mean_y=weighted.mean(price, w=ipw))
mean.t1$mean_y - mean.t0$mean_y
reg.ipw <- lm(price ~ penalty, data=data_with_CT, weights=ipw)




## regression
reg1.dat <- data_with_CT %>% filter(penalty==1)
reg1 <- lm(price ~ Q1+ Q2+ Q3 + Q4, data=data_with_CT)

reg0.dat <- data_with_CT %>% filter(penalty==0)
reg0 <- lm(price ~ Q1+ Q2+ Q3 + Q4, data=data_with_CT)

pred1_alt <- predict(reg1,new=data_with_CT)
pred0_alt <- predict(reg0,new=data_with_CT)
mean(pred1_alt-pred0_alt)




# create a table of the output variables

Estimate1 <- nn.est1$est
Estimate2<-nn.est2$est
Estimate3<-invisible(coef(reg.ipw)[2])
Estimate4<-mean(pred1_alt-pred0_alt)

output_table <- data.frame("Nearest neighbor matching with inverse variance" = Estimate1,
                           "Nearest neighbor matching with inverse variance"=Estimate2, 
                           "Inverse propensity weighting"=Estimate3,
                           "Simple Regression"= Estimate4)


knitr::kable(output_table,
             type="html", caption = "ATE", booktabs = TRUE)


```

### 8
Yes all of my estimators are very similar however my regression seems to have an estimate of zero this may because these are two different samples I am trying to draw a line from. In addition, the reason we may have similar estimates is that all these estimators are doing the same thing in terms of matching between control and treatment observations and thus yielding similar results. . 

### 9
In my opinion yes i did estimate a causal inference. Given then i can't observe a counter factual world i did my best to match the control group to the treatment group by setting their bed size into quartiles that help match similar hospitals and then see the effect on price of whether or not they underwent a penalty. 

### 10
honestly this homework was incredibly hard and frustrating. I think i spent at least 20 hours of straight coding or problem solving
Also i feel very lost in terms of the new material we are covering in class which does not help when i need to apply this knowledge with real data in the homework
one thing i did learn was i am getting better at cleaninig up the data and I believe I did good work until question 5 ish
however one thing that still frustrates me is my inability to understanding the matching function and concepts we have learned in this module.











